dev---
title: "Basic how to package"
author: "Jonas Vaclavek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

While developing web applications, CSS is used to style web application's web pages. The CSS is often written using CSS preprocessor such as SASS or LESS. 

`rless` is a package, which helps to convert LESS files to CSS files within R environment. 

This documentation is divided into three sections:

- [rless functions](#rless) section, which shows usage of the package functions
- [LESS](#less) section briefly explaining basics of LESS 
- [Using in Shiny apps](#shiny) section demonstrating `rless`intergration to Shiny application 

# rless functions {#rless}

## parse_less
The `parse_less` function is a core function of the package. It takes LESS content and converts it to CSS. 

```{r}
library(rless)

less <- "
@width: 10px;
@height: @width + 10px;

#header {
  width: @width;
  height: @height;
}
"

css <- parse_less(less)
cat(css)
```

## convert_file
Passing text into `parse_less` function is suitable for small chunks. However, for more complex codes, it is more convenient to write LESS file separately and convert the file using `convert_file` function. It requires two parameters:

* file name
* path to the file (both relative and absolute are supported)

The converted file will be stored in `tempdir` folder or use `output_folder` argument to set different output location. 

The function returns full path to the converted.

```r
less_file <- file.path("path", "to", "styles.less")
convert_file(dirname(less_file), basename(less_file), getwd())

```

## convert_folder
When having multiple LESS files, one can use `rless` function `convert_folder`. The function requires only input folder to be specified. It goes through the folder and converts all files matching a `pattern`  (*.less by default) into CSS files. `tempdir` folder is used again as a default output location. 

The function allows to specify optional argument:

* output_folder - folder, where converted files will be places
* recursive - whether to scan for files in nested folders
* pattern - file name pattern used to select files to convert

```r
styles_folder <- file.path("path", "to", "styles", "folder")

paths_to_files <- convert_folder(styles_folder)

# set different output folder
paths_to_files <- convert_folder(styles_folder, output_folder = getwd())

# convert also files in file.path(styles_folder, "nested_folder", "even_more_nested_folder")
paths_to_files <- convert_folder(styles_folder, recursive = TRUE)

# convert only files with filename ending with 'styles.less'
paths_to_files <- convert_folder(styles_folder, pattern = "^*.styles.less$")
```


List of paths to converted files is returned.

## LESS {#less}
The following sections introduce the basic concepts of LESS with a use of `rless` function `parse_less`. More information about LESS features can be found [here][less-features].

### Variables
Use variables at multiple places in your LESS/CSS.

``` {r}
library(rless)

less <- "
@width: 10px;
@height: @width + 10px;

#header {
  width: @width;
  height: @height;
}

#footer{
  width: @width;
  height: @height;
}
"

css <- parse_less(less)
cat(css)

```

### Nesting and parent selector
Basic nesting of CSS selectors
```{r}
less <- "
ul {
  li {
    color: cyan;
  }  
}
"

css <- parse_less(less)
cat(css)

```
Nesting with reference to parent.
``` {r}
less <- "
.button {
  &-ok {
    background-color: green;
  }
  && {
    backgrou-color: transparent;
  }
}
"

css <- parse_less(less)
cat(css)
```

### Mixins
Combine chunks of CSS using mixins.

``` {r}
less <- "
.bordered {
  border-top: dotted 1px black;
  border-bottom: solid 2px black;
}

#menu a {
  color: #111;
  .bordered();
}

.post a {
  color: red;
  .bordered();
}
"

css <- parse_less(less)
cat(css)

```


## Using in Shiny apps {#shiny}
One of the typical ways of creating web applications in R is using [Shiny][shiny]. This sections shows how easy is to combine it with our `rless` package.


### Single LESS file
```r
# global.R
convert_file(getwd(), "styles.less", "www")

# ui.R
shinyUI(
  fluidPage(
    tags$head(
      tags$link(rel = "stylesheet", type = "text/css", href = "styles.css")
    )
  )
)

```

### Multiple LESS files
```r
# ui.R
shinyUI(
  fluidPage(
    lapply(convert_folder(input_folder, output_folder), includeCSS)
  )
)
  
```
[shiny]: https://shiny.rstudio.com/
[less-features]: http://lesscss.org/features/
